\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{parskip}
\usepackage{array}

\lstdefinestyle{pystyle}{
  basicstyle=\ttfamily\small,
  keywordstyle=\color{blue!70!black},
  commentstyle=\color{gray},
  stringstyle=\color{orange!70!black},
  showstringspaces=false,
  breaklines=true,
  frame=single,
  language=Python,
}
\lstset{style=pystyle}

\title{DMRGPY Physics User Guide}
\author{}
\date{}

\sloppy

\begin{document}
\maketitle

This guide explains what DMRGPY computes from a \textbf{physics} point
of view --- the quantities each method returns, the formula behind
them, and when to reach for which method. It intentionally says nothing
about the solver backends (DMRG vs.\ ED, C++ vs.\ Python vs.\ Julia) or
code layout --- see \texttt{docs/documentation.tex} for that. Every
\texttt{Many\_Body\_Chain} subclass (\texttt{Spin\_Chain},
\texttt{Fermionic\_Chain}, \texttt{Bosonic\_Chain}, \ldots) exposes the
same physics-facing API described here, with \texttt{mode="DMRG"|"ED"}
selectable on almost every call so a result can be cross-checked against
an exact reference on small systems.

\tableofcontents

\section{Physical models and Hilbert spaces}

Every model is a chain of $n$ local Hilbert spaces $\mathcal
H=\bigotimes_{i=1}^n\mathcal H_i$, on which a Hamiltonian and
observables are built out of local operators.

\begin{center}
\footnotesize
\begin{tabular}{@{}p{0.28\textwidth}p{0.32\textwidth}p{0.3\textwidth}@{}}
\toprule
Chain class & Local Hilbert space & Key operators \\
\midrule
\texttt{spinchain.\allowbreak Spin\_Chain} & spin-$S$, $S\in\{\tfrac12,1,\tfrac32,2,\tfrac52,3\}$ per site & $S^x_i,S^y_i,S^z_i$ \\
\texttt{fermionchain.\allowbreak Fermionic\_Chain} & spinless fermion (occupied/empty) & $c_i,c_i^\dagger,n_i=c_i^\dagger c_i$, JW string $F_i$ \\
\texttt{fermionchain.\allowbreak Majorana\_Chain} & Majorana fermion & Majorana operators built from \texttt{Fermionic\_Chain} \\
\texttt{fermionchain.\allowbreak Spinful\_\allowbreak Fermionic\_Chain} & spin-$\tfrac12$ fermion (4 states), two interleaved spinless sites per physical site & $c_{i\sigma},c^\dagger_{i\sigma},n_{i\sigma}$; derived spin and pairing operators (see text) \\
\texttt{bosonchain.\allowbreak Bosonic\_Chain} & truncated boson Fock space, $n_i\le n_{\max}$ (default 4) & $a_i,a_i^\dagger,n_i$, occupation projectors \\
\texttt{parafermionchain.\allowbreak Parafermionic\_\allowbreak Chain} & $\mathbb Z_N$ parafermion (clock model), $N\in\{2,3,4\}$ & clock/shift operators $\sigma_i,\tau_i$, composite $\chi_i,\psi_i$ \\
\bottomrule
\end{tabular}
\normalsize
\end{center}

Spinful fermionic chains are built by \emph{interleaving} two spinless
fermionic sites per physical site (site $2i$ = spin up, site $2i+1$ =
spin down) rather than by a genuinely 4-dimensional local space, so that
the same Jordan-Wigner machinery used for spinless fermions applies
unchanged; \texttt{Spinful\_Fermionic\_Chain} wraps this bookkeeping for
you.

\section{Building a Hamiltonian and observables}

Any Hamiltonian or observable is written exactly as its
second-quantized or spin-operator expression, using operator lists
indexed by site, e.g.\ for the spin-$\tfrac12$ Heisenberg chain

\begin{equation}
H=\sum_{i=1}^{n-1}J\,\mathbf S_i\cdot\mathbf S_{i+1}=\sum_{i=1}^{n-1}J\left(S_i^xS_{i+1}^x+S_i^yS_{i+1}^y+S_i^zS_{i+1}^z\right)
\end{equation}

\begin{lstlisting}
from dmrgpy import spinchain
spins = ["S=1/2" for i in range(30)]
sc = spinchain.Spin_Chain(spins)
h = 0
for i in range(29):
    h = h + sc.Sx[i]*sc.Sx[i+1] + sc.Sy[i]*sc.Sy[i+1] + sc.Sz[i]*sc.Sz[i+1]
sc.set_hamiltonian(h)
\end{lstlisting}

Any sum of products of these operators is a valid \texttt{MultiOperator},
so essentially any local (or long-range) Hamiltonian on the given
lattice can be built term by term: single-ion anisotropy
$D(S_i^z)^2$, Dzyaloshinskii-Moriya terms $\mathbf D\cdot(\mathbf
S_i\times\mathbf S_{i+1})$, biquadratic exchange $(\mathbf S_i\cdot
\mathbf S_{i+1})^2$, hopping $t\sum_i(c_i^\dagger c_{i+1}+\text{h.c.})$,
Hubbard interaction $U\sum_i n_{i\uparrow}n_{i\downarrow}$, and so on ---
see Section~\ref{sec:cookbook} for concrete Hamiltonians.

\section{Ground-state properties}

\textbf{Ground-state energy}

\begin{equation}
E_0=\langle\mathrm{GS}|H|\mathrm{GS}\rangle=\min_{|\psi\rangle}\frac{\langle\psi|H|\psi\rangle}{\langle\psi|\psi\rangle}
\end{equation}

\begin{lstlisting}
e0 = sc.gs_energy()          # E0
wf = sc.get_gs()             # the |GS> wavefunction itself
\end{lstlisting}

\textbf{Expectation values and moments.} For any operator $O$ built the
same way as $H$,

\begin{equation}
\langle O\rangle=\langle\mathrm{GS}|O|\mathrm{GS}\rangle,\qquad
\langle O^n\rangle=\langle\mathrm{GS}|O^n|\mathrm{GS}\rangle
\end{equation}

\begin{lstlisting}
mz = [sc.vev(sc.Sz[i]).real for i in range(n)]        # local magnetization profile
e2 = sc.vev(h, npow=2)                                  # <H^2>, e.g. for fluctuations
\end{lstlisting}

\textbf{Energy fluctuation} (a measure of how sharply the DMRG/ED state
is an eigenstate, and physically the variance of $H$ in the prepared
state):

\begin{equation}
\delta E=\sqrt{\langle H^2\rangle-\langle H\rangle^2}
\end{equation}

\begin{lstlisting}
de = sc.gs_energy_fluctuation()
\end{lstlisting}

\textbf{Static two-point correlators.} \texttt{sc.vev(sc.Sz[0]*sc.Sz[i])}
gives $\langle S^z_0 S^z_i\rangle$ directly; \texttt{correlator.get\_correlator}
provides shorthand names for common correlators over a list of site
pairs, e.g.\ \texttt{"SS"} for the full dot product

\begin{equation}
\langle\mathbf S_i\cdot\mathbf S_j\rangle=\langle S_i^xS_j^x\rangle+\langle S_i^yS_j^y\rangle+\langle S_i^zS_j^z\rangle
\end{equation}

or fermionic correlators like \texttt{"cdc"} ($\langle c_i^\dagger
c_j\rangle$), \texttt{"density"}/\texttt{"densitydensity"} ($\langle
n_in_j\rangle$), and pairing correlators \texttt{"delta"}/
\texttt{"deltadeltad"}. These static correlators are the equal-time
limit of the dynamical correlators in Section~\ref{sec:dynamical}, and
are what you Fourier-transform to get an equal-time structure factor
$S(q)=\sum_{i,j}e^{iq(i-j)}\langle O_iO_j\rangle$.

\section{Excited states and energy gaps}

\textbf{Low-lying spectrum.}

\begin{lstlisting}
es = sc.get_excited(n=4)          # [E0, E1, E2, E3]
gap = sc.get_gap()                # E1 - E0
\end{lstlisting}

\texttt{get\_excited\_states(n, purify=True)} additionally returns the
wavefunctions; \texttt{purify=True} re-diagonalizes $H$ in the
Gram-Schmidt orthogonalized subspace spanned by the raw excited-state
MPS, correcting for near-degenerate states that DMRG's iterative solvers
can otherwise mix.

\textbf{Sector (charge) gaps.} For a conserved quantity $A$ with
$[H,A]=0$ (e.g.\ total particle number $\hat N=\sum_i n_i$, or total
$S^z$), the gap to the lowest state with $\langle A\rangle$ shifted by
$d$ from the ground-state sector is obtained by adding a
Lagrange-multiplier penalty and increasing $\lambda$ until the
constraint is satisfied:

\begin{equation}
H_\lambda=H+\lambda\big(A-\langle A\rangle_0-d\big)^2,\qquad
\Delta_A(d)=E_\lambda-E_0
\end{equation}

\begin{lstlisting}
gap_charge = fc.get_charge_gap(d=2)   # gap to add/remove a pair of particles
\end{lstlisting}

This is exactly how a superconducting/charging gap is extracted in a
finite fermionic chain: $d=2$ probes the energy cost of adding a pair of
particles (Cooper-pair-like excitation) rather than a single particle,
which avoids odd/even-parity finite-size artifacts.

\textbf{Example: the Haldane gap.} For the spin-1 Heisenberg chain, the
famous Haldane gap between the (nearly four-fold degenerate, due to
fractionalized $S=\tfrac12$ edge states) ground-state manifold and the
first bulk excitation is

\begin{lstlisting}
es = sc.get_excited(n=6)
haldane_gap = es[4] - es[0]     # skip the 4 near-degenerate edge states
\end{lstlisting}

\textbf{Non-Hermitian Hamiltonians.} For $H\neq H^\dagger$ (complex
hopping amplitudes, gain/loss terms, PT-symmetric models, \ldots) the
``ground state'' convention throughout dmrgpy is the eigenvalue with the
smallest real part. Eigenvalues come in left/right eigenpairs,
\begin{equation*}
H\,|\psi_R\rangle=E\,|\psi_R\rangle,\qquad
H^\dagger|\psi_L\rangle=E^{*}|\psi_L\rangle,\qquad
\langle\psi_L|\psi_R\rangle=1 .
\end{equation*}
On every session backend (ITensor v3, ITensor v2, and the pure-Python
engine --- \lstinline|itensor_version| 2, 3 or \lstinline|"python"|),
\lstinline|gs_energy()| solves this with a genuine non-Hermitian DMRG
(NH-DMRG) --- a port of
ITensorNHDMRG.jl\footnote{\url{https://github.com/tipfom/ITensorNHDMRG.jl}}
in its default configuration: independent Arnoldi solves of the two-site
eigenproblem for $H$ and $H^\dagger$ (``onesided'' solver, targeting the
smallest real part), with both MPS truncated by the same isometry
obtained from the Hermitian average $\rho=(\rho_L+\rho_R)/2$ of the
left/right reduced density matrices (the ``fidelity'' algorithm of
Yamamoto \emph{et al.}, PRB \textbf{105}, 205125 (2022)). The full
eigenpair is available directly:

\begin{lstlisting}
e, psil, psir = sc.nhdmrg()     # E, left and right eigenvector MPS
sc.gs_energy()                  # same E; stores psir as the ground state
\end{lstlisting}

Because a non-Hermitian ``energy'' is not a variational bound,
\lstinline|nhdmrg()| certifies convergence through both eigen-residuals
$\lVert H|\psi_R\rangle-E|\psi_R\rangle\rVert$ and
$\lVert H^\dagger|\psi_L\rangle-E^{*}|\psi_L\rangle\rVert$ and re-runs
from a fresh random state (up to \lstinline|ntries| times) if either
stalls; \lstinline|krylovdim|/\lstinline|restarts| tune the per-bond
Arnoldi effort. The MPS Arnoldi route (\lstinline|get_excited_states|,
now used for non-Hermitian \emph{excited} states, $n\ge 2$) remains
available on every backend, but is typically several orders of magnitude
less accurate than NH-DMRG at comparable cost --- see
\texttt{examples/non\_hermitian/nhdmrg\_VS\_ED\_VS\_arnoldi}, which
cross-checks NH-DMRG on all three backends against exact diagonalization
and the Arnoldi route on an interacting fermionic chain with a staggered
imaginary potential.

\section{Entanglement and quantum information}

\textbf{Entanglement entropy of a real-space bipartition.} Cutting the
chain at bond $(i,i+1)$, the Schmidt decomposition of the ground state
is $|\mathrm{GS}\rangle=\sum_\alpha\lambda_\alpha|\alpha\rangle_L\otimes
|\alpha\rangle_R$ and the von Neumann entanglement entropy of either
half is

\begin{equation}
S_i=-\sum_\alpha\lambda_\alpha^2\log\lambda_\alpha^2=-\mathrm{Tr}\big[\rho_L\log\rho_L\big],\qquad
\rho_L=\mathrm{Tr}_R|\mathrm{GS}\rangle\langle\mathrm{GS}|
\end{equation}

\begin{lstlisting}
s = wf.get_bond_entropy(i, i+1)
\end{lstlisting}

\texttt{get\_site\_entropy(i)}/\texttt{get\_pair\_entropy(i,j)} instead
build a reduced density matrix from local \emph{projectors} (e.g.\
$S^z=\pm\tfrac12$ for spin-$\tfrac12$, occupation number for fermions)
rather than a full MPS bond cut, which lets you ask about the
entanglement of a single site or a pair of (possibly non-adjacent)
sites with the rest of the system.

\textbf{Mutual information} between two sites (or subsystems) $i,j$:

\begin{equation}
I(i,j)=S_i+S_j-S_{ij}
\end{equation}

\begin{lstlisting}
mi = wf.get_mutual_information(1, 2)
\end{lstlisting}

quantifies total (classical + quantum) correlation between $i$ and $j$,
and is often used to map out effective couplings or emergent degrees of
freedom (e.g.\ between the two edge spins of a Haldane chain).

\textbf{CFT central charge.} For a critical (gapless) chain, the
entanglement entropy of a cut at position $\ell$ in an open chain of
length $L$ follows the Calabrese-Cardy formula:

\begin{equation}
S(\ell)=\frac{c}{6}\log\!\left[\frac{2L}{\pi a}\sin\frac{\pi\ell}{L}\right]+\text{const.}
\end{equation}

DMRGPY computes $S(\ell)$ at every bond and least-squares fits this
formula to extract the central charge $c$ --- the universal number that
identifies the underlying conformal field theory (e.g.\ $c=\tfrac12$
for the critical Ising chain, $c=1$ for a free boson / XX chain):

\begin{lstlisting}
c = wf.get_CFT_central_charge()
\end{lstlisting}

\textbf{Single-particle correlation matrix and orbital entanglement.}
For fermionic systems, the correlation matrix

\begin{equation}
C_{ij}=\langle c_i^\dagger c_j\rangle
\end{equation}

is directly accessible (\texttt{sc.get\_correlation\_matrix()}), and
diagonalizing it, $C=U\,n\,U^\dagger$, gives natural-orbital occupations
$n_\alpha\in[0,1]$ (\texttt{get\_correlated\_orbitals}) and an
orbital-resolved entanglement entropy

\begin{equation}
S=-\sum_\alpha\Big[n_\alpha\log n_\alpha+(1-n_\alpha)\log(1-n_\alpha)\Big]
\end{equation}

(\texttt{get\_correlation\_entropy}). \texttt{get\_highorder\_correlation\_matrix}
and \texttt{get\_four\_correlation\_tensor} give the corresponding
two-particle correlators $\langle c_i^\dagger c_j^\dagger c_l
c_k\rangle$ and $\langle c_i^\dagger c_j c_k^\dagger c_l\rangle$.

\section{Dynamical (frequency-dependent) correlators}
\label{sec:dynamical}

The central quantity is a retarded correlator (Green's function)
between two operators $A,B$,

\begin{equation}
G_{AB}(\omega)=\langle\mathrm{GS}|A\,\frac{1}{\omega-H+E_0+i\delta}\,B|\mathrm{GS}\rangle,\qquad
S_{AB}(\omega)=-\frac1\pi\,\mathrm{Im}\,G_{AB}(\omega)
\end{equation}

where $\delta$ is a small broadening. Choosing $A=B=S^z_i$ at the same
site gives the local dynamical spin structure factor
$S^{zz}_{ii}(\omega)$ (what a local probe like NMR/ESR couples to);
choosing $A=S^z_i$, $B=S^z_j$ at different sites and Fourier-transforming
over $i-j$ gives the momentum-resolved dynamical structure factor
$S(q,\omega)$ measured in inelastic neutron scattering. All five
submodes below compute $G_{AB}(\omega)$ (or equivalently
$S_{AB}(\omega)$); they differ only in \emph{how}, and therefore in what
energy range/resolution/cost trade-off they offer:

\begin{lstlisting}
(x, y) = sc.get_dynamical_correlator(mode="DMRG", submode="KPM",
                                      name=(sc.Sz[0], sc.Sz[0]))
\end{lstlisting}

\textbf{\texttt{submode="KPM"} --- Kernel Polynomial Method.} $H$ is
linearly rescaled into $\tilde H=(H-b)/a$ so its full spectrum lies in
$[-1,1]$ (\texttt{kpm\_scale}, default 0.7, sets how much margin is left
inside $[-1,1]$), then $G_{AB}$'s spectral function is expanded in
Chebyshev polynomials $T_m(\tilde H)$:

\begin{equation}
\mu_m=\langle\mathrm{GS}|A\,T_m(\tilde H)\,B|\mathrm{GS}\rangle,\qquad
S_{AB}(\tilde\omega)=\frac{1}{\pi\sqrt{1-\tilde\omega^2}}\left[\mu_0+2\sum_{m=1}^{N-1}g_m\,\mu_m\,T_m(\tilde\omega)\right]
\end{equation}

with Jackson-kernel damping coefficients $g_m$ suppressing Gibbs
ringing from the truncation at $N$ moments (\texttt{kpm\_n\_scale}
scales $N$, i.e.\ the energy resolution, relative to the rescaled
bandwidth). This is the default, general-purpose method: robust, works
across the whole spectrum at once, cost grows only linearly with the
number of moments.
The band edges entering the rescaling are obtained variationally: the
lower edge reuses the ground-state energy, and the upper edge runs a
deliberately reduced-effort DMRG on $-H$ (few sweeps at modest bond
dimension) --- it is only a spectral \emph{bound}, protected by the
\texttt{kpm\_scale} margin, not a physical result, and a variational
underestimate only shrinks the number of moments. If the bound is ever
too tight for the chosen \texttt{kpm\_scale}, the moment recursion
detects the resulting exponential divergence and aborts with an
explicit error rather than returning a silently wrong spectrum.

\textbf{\texttt{submode="CVM"} --- correction-vector method.} Instead of
a global polynomial expansion, this solves directly for the correction
vector at one frequency $\omega$ at a time, via the positive-definite
linear system

\begin{equation}
\big[(H-\omega-E_0)^2+\eta^2\big]\,x_c=-\eta\,B|\mathrm{GS}\rangle
\end{equation}

(solved by conjugate gradient in MPS form), from which

\begin{equation}
x=(\omega+E_0+i\eta-H)^{-1}B|\mathrm{GS}\rangle=i\,x_c+\frac{H-\omega-E_0}{\eta}\,x_c,\qquad
G_{AB}(\omega)=\langle\mathrm{GS}|A|x\rangle
\end{equation}

Here $\eta$ (\texttt{delta}) is the artificial broadening that
regularizes the resolvent at a real frequency. CVM is more accurate at a
single targeted frequency/energy window (e.g.\ zooming in on a sharp
resonance) than a global KPM expansion, at the cost of re-solving the
linear system for every $\omega$ on the requested grid. Two things keep
that per-$\omega$ cost down. First, everything $\omega$-independent is
computed once per sweep instead of once per point --- in particular the
right-hand side $-\eta B|\mathrm{GS}\rangle$. Second, the CG loop
terminates early once the bond-dimension truncation
(\texttt{cvm\_maxm}) puts a floor under the reachable residual ---
past that floor the iteration cannot improve the tracked best solution
any further (the truncated recurrence in fact diverges), so the solver
returns the best correction vector seen instead of burning the full
\texttt{cvm\_nit} iteration budget (\texttt{cvm\_patience} sets how
many iterations without meaningful improvement conclude the floor is
reached, and \texttt{cvm\_blowup} how far past the best residual the
running one may diverge before stopping). Neither feature changes the
answer.
Each point reports its CG iteration count and best residual; if that
residual stalls far above \texttt{cvm\_tol}, the correction vector is
not converged at this bond dimension and the fix is a larger
\texttt{cvm\_maxm}, not more iterations. (Warm-starting each point's CG
from the neighboring point's correction vector was tried and measured
to \emph{hurt} --- truncated CG from a nearby-but-wrong start can
stagnate at a much worse residual than from the cold start --- so each
point is solved independently.)

\textbf{\texttt{submode="TD"} --- time-dependent DMRG.} Real-time
evolution gives the correlator directly in the time domain,

\begin{equation}
C(t)=\langle\mathrm{GS}|A(t)B(0)|\mathrm{GS}\rangle=e^{iE_0t}\langle\mathrm{GS}|A\,e^{-iHt}\,B|\mathrm{GS}\rangle
\end{equation}

which is then windowed with an exponential damping factor $e^{-\delta
t}$ (equivalent to a Lorentzian broadening of width $\delta$ in
frequency) and Fourier transformed,

\begin{equation}
S_{AB}(\omega)=\frac1\pi\,\mathrm{Re}\!\int_0^{T}\!dt\;e^{i\omega t}\,C(t)\,e^{-\delta t}
\end{equation}

The total simulated time $T$ (\texttt{damping\_periods}/$\delta$) must
be long enough that the damping has suppressed truncation ringing by
$t=T$. Frequency resolution is set by $T$ (via the usual $\Delta\omega
\sim1/T$ time-frequency uncertainty), so this method is best when you
want fine resolution over a \emph{narrow} frequency window, at the cost
of a real dynamical simulation (bond dimension grows with entanglement
generated during the evolution).

\textbf{\texttt{submode="TDZ"} --- complex-time evolution (Cao, Lu,
Stoudenmire \& Parcollet, arXiv:2311.10909).} Real-time evolution
(\texttt{"TD"} above) grows entanglement, so the MPS bond dimension
needed for a given accuracy grows with the simulated time $T$. TDZ
instead evolves along a complex time contour

\begin{equation}
z(t,\alpha_0)=\int_0^t e^{-i\alpha_0 f(t')}\,dt',\qquad f(t)=e^{-t\omega_0},\qquad \omega_0=2\pi/t_{\max}
\end{equation}

Since $\mathrm{Im}\,z(t,\alpha_0)<0$ for $\alpha_0>0$, this
progressively damps high-energy content as it evolves, so the bond
dimension needed for a given accuracy grows far more slowly than under
real-time evolution alone (the original paper reports $\chi\sim20$--$30$
vs $\chi\sim500$--$700$ for comparable accuracy on the Anderson impurity
model). The true real-time ($\alpha_0=0$) correlator is then recovered
order by order via a perturbative Taylor expansion in $\alpha_0$ around
the simulated contour,

\begin{equation}
C(t,0)\approx\phi^{(0)}(t,\alpha_0)+\sum_{n=1}^{n_{\max}}g^{(n)}(t,\alpha_0)
\end{equation}

where $\phi^{(n)}(t,\alpha_0)=\langle H^n B|\mathrm{GS}\rangle\cdot
|\psi(t,\alpha_0)\rangle$ (precomputed once per $n$, reused as a fixed
overlap target at every time step) and $g^{(n)}$ are explicit
combinatorial expressions in $\phi^{(1..n)}$ and the pure contour
integrals $J^{(n)}(t,\alpha_0)=-i\,\partial^n_{\alpha_0}z(t,\alpha_0)$
(see the paper's Appendix B; this implementation hardcodes $n\le4$,
which the paper finds always suffices for $\alpha_0\lesssim0.3$). The
reconstructed $C(t,0)$ is then windowed/Fourier-transformed exactly as
in \texttt{"TD"}.

\begin{verbatim}
(x, y) = sc.get_dynamical_correlator(mode="DMRG", submode="TDZ",
                                      name=(sc.Sz[0], sc.Sz[0]),
                                      alpha0=0.1, n_max=4, dt=0.05)
\end{verbatim}

\texttt{alpha0} is the contour angle parameter (larger reduces the bond
dimension needed further, but requires a larger \texttt{n\_max} to
reconstruct the real axis accurately); \texttt{n\_max} ($\le4$) is the
reconstruction order; \texttt{dt}/\texttt{tmax}/\texttt{nt} set the
underlying time step/duration exactly as in \texttt{"TD"}. Uses
two-site TDVP when available (\texttt{itensor\_version} 3 or
\texttt{"python"}, \texttt{tevol\_method="TDVP"}, the paper's own setup)
or falls back to the MPO-Taylor propagator otherwise
(\texttt{itensor\_version=2}, which has no TDVP) --- the same
TDVP-vs-Taylor choice \texttt{"TD"} already makes. Current scope: only
the ``greater'' branch of the correlator is computed (the same
simplification \texttt{"TD"} itself already makes), so this is best
used the same way as \texttt{"TD"}: high-resolution work in a narrow
frequency window, now reachable at a lower bond-dimension cost for a
given simulated time.

\textbf{\texttt{submode="EX"} --- exact diagonalization in a truncated
DMRG subspace.} Builds $A$, $B$, $H$ explicitly in the subspace spanned
by the lowest \texttt{nex} (orthogonalized) DMRG excited states, then
evaluates the exact Lehmann sum in that subspace,

\begin{equation}
G_{AB}(\omega)=\sum_{n=1}^{n_{ex}}\frac{\langle\mathrm{GS}|A|n\rangle\langle n|B|\mathrm{GS}\rangle}{\omega-E_n+i\delta}
\end{equation}

i.e.\ a small, explicit sum over poles at the computed excited-state
energies $E_n$, each with residue given by the transition matrix
elements. Cheap and exact \emph{within} the truncated subspace; only as
good as how many/which excited states were computed.

\textbf{\texttt{submode="maxent"} --- maximum-entropy reconstruction.}
Reconstructs a positive-definite spectral function from a finite set of
moments $\langle(H-E_0)^k\rangle$ using a maximum-entropy method
(\texttt{distribution.get\_distribution\_maxent}), rather than a
Chebyshev expansion --- useful when positivity of the reconstructed
$S(\omega)$ matters more than matching KPM's polynomial-expansion
artifacts.

\textbf{Choosing a method:} KPM (default) for a first look at the full
spectrum; CVM or TD when you need high resolution in a specific,
narrow frequency window; TDZ instead of TD when that window also needs
long simulated times/low frequencies, where TD's real-time bond-dimension
growth becomes limiting; EX when a handful of excited states already
capture the physics (e.g.\ a small gapped system); maxent when you want
a guaranteed-positive reconstruction from limited moment data (e.g.\
combined with finite-temperature ED, see Section~\ref{sec:thermal}).

\section{Real-time dynamics: quenches}

Beyond frequency-domain correlators, DMRGPY directly simulates real-time
unitary evolution $|\psi(t)\rangle=e^{-iHt}|\psi(0)\rangle$ and measures
an observable along the way:

\begin{lstlisting}
from dmrgpy import timedependent
wf0 = sc.get_gs()                          # prepare |GS> of H0
sc.set_hamiltonian(h1)                     # quench to a different H
(ts, sz) = timedependent.evolve_and_measure(sc, operator=sc.Sz[0],
                                             nt=200, dt=1e-2, wf=wf0)
\end{lstlisting}

giving $\langle\psi(t)|S^z_0|\psi(t)\rangle$ as a function of time after
a Hamiltonian quench $H_0\to H_1$ --- the standard non-equilibrium
quantum quench setup: prepare the ground state of one Hamiltonian
(e.g.\ a symmetry-broken/N\'eel-ordered $H_0$), then let it evolve
under a different $H_1$ (e.g.\ the isotropic Heisenberg point) and watch
observables relax, oscillate, or thermalize. \texttt{evolution\_ABA}
similarly lets you apply an operator $A$ as an instantaneous local
quench (e.g.\ flip a spin, or add a particle) before evolving and
measuring $B(t)$. \texttt{timeevolution.imaginary\_exponential} computes
autocorrelation functions directly, $\langle\psi_0|e^{iHt}|\psi_0
\rangle$, without a separate measurement operator.

\section{Density of states}

\textbf{Many-body density of states.} The full many-body spectral
density

\begin{equation}
\rho(E)=\sum_n\delta(E-E_n)
\end{equation}

(sum over \emph{all} eigenstates of $H$, not resolved by any operator)
is obtained via the same KPM machinery as Section~\ref{sec:dynamical}
applied directly to $H$ itself:

\begin{lstlisting}
from dmrgpy import dos
es, rho = dos.get_dos(sc)
\end{lstlisting}

\textbf{Single-particle local density of states.} For fermionic chains,
the physically distinct quantity usually meant by ``DOS'' (e.g.\ as
measured by scanning tunneling spectroscopy) is the local
single-particle spectral function at site $i$,

\begin{equation}
A_i(\omega)=-\frac1\pi\,\mathrm{Im}\,G^R_{ii}(\omega)
\end{equation}

with $G^R_{ii}(\omega)$ built from $\langle c_i(t)c_i^\dagger(0)\rangle$
($\omega>0$) and $\langle c_i^\dagger(t)c_i(0)\rangle$ ($\omega<0$),
i.e.\ the particle-addition and particle-removal (electron/hole)
branches of the local Green's function concatenated across $\omega=0$.
There is no ready-made helper for this (the module this section used to
point to, \texttt{fermiondos.py}, relied on a
\texttt{get\_dynamical\_correlator} calling convention that no longer
exists and was removed) --- build the two branches directly with
\texttt{get\_dynamical\_correlator} instead, using
\texttt{(fc.C[i], fc.Cdag[i])} for the particle-removal branch and
\texttt{(fc.Cdag[i], fc.C[i])} for the particle-addition branch, then
concatenate.

\section{Finite temperature}
\label{sec:thermal}

Thermal (mixed-state, finite-$T$) expectation values are obtained via
\textbf{purification}: each physical site is paired with an ancilla
site, and the maximally-entangled state of every physical-ancilla pair
(e.g.\ the singlet-forming Heisenberg coupling $\mathbf S_i^{\rm
phys}\cdot\mathbf S_i^{\rm anc}$ as the ``Hamiltonian'' preparing that
state) is exactly the purification of the infinite-temperature
($T=\infty$) physical density matrix $\rho\propto\mathbb 1$.
Imaginary-time evolving this purified state under the \emph{physical}
Hamiltonian $H$,

\begin{equation}
|\Psi(\beta)\rangle=e^{-\beta H/2}|\Psi(0)\rangle,\qquad\beta=1/T
\end{equation}

(stepped via repeated small first-order updates
$|\Psi(\beta+\Delta\beta)\rangle\propto(1-\Delta\beta\,H)|\Psi(\beta)\rangle$,
renormalized at each step) and tracing out the ancillas gives exactly
the thermal (Gibbs) density matrix of the physical chain,
$\rho(T)=\mathrm{Tr}_{\rm anc}|\Psi(\beta)\rangle\langle\Psi(\beta)|
\propto e^{-\beta H}$:

\begin{lstlisting}
from dmrgpy import thermal
tc = thermal.Thermal_Spin_Chain(spins, T=0.1)
\end{lstlisting}

$T=0$ recovers ordinary ground-state DMRG. For small systems
($n\lesssim14$), \texttt{get\_correlation\_matrix(T=...)} instead
computes the exact thermal average directly by brute-force ED,
$\rho=\sum_nZ^{-1}e^{-E_n/T}|n\rangle\langle n|$ --- a useful
cross-check of the purification approach.

\section{Topological invariants}

\textbf{Many-body Berry phase.} For a ground state that depends on an
adiabatic parameter $k$ threaded around a closed loop (e.g.\ inserted
flux), the discretized Berry phase

\begin{equation}
\gamma=\arg\prod_{k}\langle\psi(k)|\psi(k+\delta k)\rangle
\end{equation}

quantized (typically in units of $\pi$) signals a topologically
nontrivial ground state / obstruction to adiabatic continuity --- the
many-body generalization of a Zak/Berry phase, computed by running DMRG
at a discrete set of parameter points and chaining ground-state
overlaps.

\textbf{Single-particle Berry phase / Wilson loop.} For a
translationally invariant single-particle Hamiltonian $H(k)$ with a set
of occupied bands, the non-Abelian Wilson loop over one Brillouin-zone
circuit is

\begin{equation}
W=\det\Big[\textstyle\prod_k U(k,k+\delta k)\Big],\qquad
U_{mn}(k,k+\delta k)=\langle u_m(k)|u_n(k+\delta k)\rangle
\end{equation}

with $m,n$ running over occupied bands; $\gamma=\arg W$ is the
polarization/Zak phase of that band manifold (and, combined with a scan
over a second momentum direction, the ingredient for a Chern number).

\section{Mean-field decoupling}

For a spin Hamiltonian with exchange couplings $J_{ij}$, a
self-consistent mean-field (Weiss-field) decoupling replaces the
two-body exchange term with a one-body field,

\begin{equation}
H_{\rm MF}=\sum_i\mathbf h_i\cdot\mathbf S_i,\qquad
\mathbf h_i=(1-p)\sum_jJ_{ij}\langle\mathbf S_j\rangle
\end{equation}

iterated to self-consistency: solve for $\langle\mathbf S_i\rangle$
under $H_{\rm MF}$ (by DMRG), rebuild $\mathbf h_i$ from the new
expectation values, mix old/new fields, and repeat until
$\max_i|\Delta\langle\mathbf S_i\rangle|$ falls below a tolerance. The
parameter $p\in[0,1]$ interpolates between pure mean-field theory
($p=0$) and keeping a fraction $p$ of the original many-body exchange
treated exactly alongside the self-consistent field (useful for hybrid
mean-field-plus-fluctuations treatments of, e.g., magnetically ordered
phases where pure MF overestimates order).

\begin{lstlisting}
from dmrgpy import meanfield
meanfield.spinchain_meanfield(sc, p=0.0)
\end{lstlisting}

\section{Fidelity susceptibility and quantum phase transitions}

For a Hamiltonian $H(\lambda)=H_0+\lambda H_1$ depending on a tuning
parameter $\lambda$, the fidelity susceptibility measures how sharply
the ground state changes as $\lambda$ varies --- and diverges at a
quantum phase transition, where the gap to the first excited state
closes:

\begin{equation}
\chi(\lambda)=\sum_{n\neq0}\frac{|\langle 0|H_1|n\rangle|^2}{(E_n-E_0)^2+\delta^2}
\qquad(\text{perturbative form, }\texttt{fmode="PT"})
\end{equation}

\texttt{get\_fidelity}'s \textbf{default} is not this perturbative form
but a non-perturbative estimator (\texttt{fmode="derivative"}) that
computes $\chi$ directly from finite differences of the ground-state
overlap matrix between nearby $\lambda$ values, which also handles a
(near-)degenerate ground-state manifold via a smooth gauge choice:

\begin{lstlisting}
from dmrgpy import fidelity
chi = fidelity.get_fidelity(sc, h0, h1, lam, n=3) # fmode="derivative" (default)
chi_pt = fidelity.get_fidelity(sc, h0, h1, lam, n=3, fmode="PT") # perturbative form above
\end{lstlisting}

Scanning $\lambda$ and plotting $\chi(\lambda)$, a peak (sharpening and
diverging with system size) locates a quantum critical point without
needing to know its universality class in advance.

\section{Ground-state degeneracy}

Exact and near (e.g.\ symmetry-protected, or finite-size-split
topological) ground-state degeneracies are estimated by a narrow,
super-Gaussian-broadened level count around the ground-state energy
(\texttt{degeneracy.py}'s \texttt{gs\_degeneracy\_simple}/
\texttt{eigenvalue\_degeneracy}):

\begin{equation}
g(E_0)\approx\sum_i\exp\!\left[-\left(\frac{(E_i-E_0)^2}{\delta}\right)^2\right]
\end{equation}

Note the quartic falloff in $(E_i-E_0)$ (not a plain Gaussian) --- this
makes the window considerably narrower than $\delta$ would suggest for a
standard Gaussian, so \texttt{delta} needs to be picked accordingly
(typically larger than the target energy resolution) to count
near-degenerate levels rather than only exactly-degenerate ones.

summed over a growing number of low-lying computed eigenstates $E_i$
until the count converges --- a value near an integer $g$ signals a
$g$-fold degenerate (or near-degenerate, at the working precision
$\delta$) ground-state manifold, as expected e.g.\ for the four
edge-state-split ground states of an open Haldane chain, or a
symmetry-broken ordered phase.

\begin{lstlisting}
from dmrgpy import degeneracy
g = sc.get_gs_degeneracy()
\end{lstlisting}

\section{Reduced density matrices and operator distributions}

\textbf{Reduced density matrix.} The single-site reduced density matrix
$\rho_i=\mathrm{Tr}_{j\neq i}|\mathrm{GS}\rangle\langle\mathrm{GS}|$ is
available directly (\texttt{sc.get\_rdm(i=0)}), the basic object
entanglement entropies, local observables, and further post-processing
(e.g.\ local susceptibilities) are built from.

\textbf{Operator distributions.} More generally, the full probability
distribution of an arbitrary Hermitian operator $X$ (not just its
expectation value) in the ground state,

\begin{equation}
P(x)=\langle\mathrm{GS}|\,\delta(X-x)\,|\mathrm{GS}\rangle
\end{equation}

is computed via the same KPM machinery as the dynamical correlators of
Section~\ref{sec:dynamical} (\texttt{sc.get\_distribution}), or
reconstructed from a finite set of raw moments $\langle X^k\rangle$ via
maximum entropy (\texttt{get\_distribution\_maxent}). Useful for e.g.\
full counting statistics of a conserved charge, or distinguishing a
sharply peaked (well-defined quantum number) ground state from a
broadly spread one.

\section{Post-processing tools}

\begin{itemize}
  \item \textbf{Analytic continuation} (\texttt{analyticcontinuation.py}):
    Pad\'e continuation of a correlator known on the
    imaginary/complex-frequency axis (e.g.\ from a Matsubara-like or
    complex-shifted CVM calculation, Section~\ref{sec:dynamical}'s
    \texttt{submode="CVMimag"}) to the real frequency axis, where the
    physical spectral function lives.
  \item \textbf{Function fitting} (\texttt{functionfit.py}): a generic
    multi-start Powell minimizer used e.g.\ to fit the Calabrese-Cardy
    entropy formula in Section~4.
  \item \textbf{Finite-size extrapolation} (\texttt{extrapolate.py}):
    polynomial extrapolation in $1/L$ of a size-dependent quantity
    $y(L)$ toward the thermodynamic limit $L\to\infty$ --- standard
    practice for extracting bulk quantities (energy density, order
    parameters, gaps) from finite DMRG chains.
  \item \textbf{Maximum-entropy reconstruction} (\texttt{reconstruct.py}):
    reconstructs a positive spectral function from a truncated moment
    expansion, underlying both the \texttt{"maxent"} dynamical-correlator
    submode and \texttt{get\_distribution\_maxent}.
\end{itemize}

\section{Worked-example cookbook}
\label{sec:cookbook}

\textbf{Central charge of a critical transverse-field Ising chain}
($H=\sum_iS_i^zS_{i+1}^z+\tfrac12\sum_iS_i^x$ at the critical field,
expected $c=\tfrac12$):

\begin{lstlisting}
sc.maxm = 200          # larger bond dimension: needed at criticality
wf = sc.get_gs()
print(wf.get_CFT_central_charge())
\end{lstlisting}

\textbf{Haldane gap of the spin-1 chain} (see Section~4):

\begin{lstlisting}
spins = ["S=1" for i in range(n)]
sc = spinchain.Spin_Chain(spins)
# ... build Heisenberg h ...
es = sc.get_excited(n=6)
print("Haldane gap:", es[4]-es[0])
\end{lstlisting}

\textbf{Charge gap of a Hubbard chain} (see Section~4):

\begin{lstlisting}
print("Single-particle gap:", fc.get_gap())
print("Charge (pair) gap:", fc.get_charge_gap(d=2))
\end{lstlisting}

\textbf{Fidelity susceptibility across the Ising transition} (see
Section~12):

\begin{lstlisting}
h0 = 4*sum(sc.Sz[i]*sc.Sz[i+1] for i in range(n-1))   # Ising coupling
h1 = 2*sum(sc.Sx[i] for i in range(n))                # transverse field
for lam in lambdas:
    chis.append(fidelity.get_fidelity(sc, h0, h1, lam, n=3))
\end{lstlisting}

\textbf{Momentum- and frequency-resolved dynamical structure factor}
$S(q,\omega)$ (see Section~\ref{sec:dynamical}), by combining the
site-resolved KPM correlator with a lattice Fourier transform:

\begin{lstlisting}
Sqw = {}
for i in range(n):
    for j in range(n):
        x, y = sc.get_dynamical_correlator(submode="KPM", name=(sc.Sz[i], sc.Sz[j]))
        Sqw[(i, j)] = (x, y)          # combine with sum_ij e^{iq(i-j)} S_ij(w) offline
\end{lstlisting}

\end{document}
